---
title: "Data Cleaning in R"
author: "Taneja Young"
output: "word_document"
---

####Import file and name as data_frame

```{r}
PAR_ParkingTxns <- read.csv("P:/Capstone/PAR_ParkingTxns.csv")
data_frame = PAR_ParkingTxns
```

####Check out stats on file.
```{r}
str(data_frame)
View(summary(data_frame))
```

####Attach file.
```{r}
attach(data_frame)
```

####Load package DPLYR. This is for the rbind_all function, and also for filtering later on.
```{r}
install.packages("dplyr")
library(dplyr)
```


```{r}
# table_stats1 will show the number of elements, unique elements and NAs
# in each column of a data table.
# table_stats2 will list all unique values and show the frequency of 
# each unique value in the column.

data_frame_stats1 = data.frame()
data_frame_stats2 = data.frame()

#Let x be the column that you are interested in. 

for (i in (1:ncol(data_frame)))
{
  x = data_frame[,i]

  NumElements = length(x)
  NumUniqueElements = length(unique(x))
  UniqueElements = unique(x)
  ElementFrequency = data.frame(table(x))
  ElementFrequencyFrequency = data.frame(table(ElementFrequency$Freq))
  NumNAs = sum(is.na(x))

  y = data.frame((colnames(data_frame)[i]),NumElements,NumUniqueElements,NumNAs)
  data_frame_stats1 = rbind(data_frame_stats1,y)
  
  z = data.frame((colnames(data_frame)[i]),ElementFrequency)
  data_frame_stats2 = bind_rows(data_frame_stats2,z)
}

View(data_frame_stats1)
View(data_frame_stats2)
```  

**PaymentNumber** 
There are 174288 NA values.PaymentNumber is not always unique, but there are 168052 unique values out of a total of 364638 transactions. 
+166809 of these unique values occur only 1 time
+1239 values occur 2 times
+1 value occurs 3 times
+1 value occurs 4 times
+1 value - zero (0) - occurs 21056 times 
>What does a zero payment value mean?


**TicketType** 
There are 7 unique ticket types: 1, 2, 4, 257, 258, 259.

**EntryDeviceNumber**  
There are 5 Entry Device Numbers: 101, 102, 301, 302, 701.
>There are some zero's for EntryDeviceNumber. What are these?

**CreditCardNumber**
There are only 860 unique **CreditCardNumber**s. 
>Does this mean that there are only 860 unique visitors to the parking garage?
There are 176193 NA values under **CreditCardNumber**.
Notice that all ticket types where there are credit card numbers, are TicketType 2.

If there is **EntryDevice**==0, then there **EntryDateTime**==default, and if there is **PaymentNumber**==0 or ==NA, then **LastPayDate**==default. Therefore, it is clear that **EntryDateTime** is triggered by **EntryDevice**, and **LastPayDate** is triggered by **PaymentNumber**.


```{r}
View(table(TicketType,ExitDeviceNumber))
```
It seems that 301 and 302 are credit card only devices, whereas
201, 202 and 203 take any payment method.

Below, we try to determine whether there is a relationship between where someone enters and where (s)he leaves.
```{r}
heat.data = table(EntryDeviceNumber,ExitDeviceNumber)
View(table(EntryDeviceNumber,CreditCardNumber,ExitDeviceNumber))
```

####Now that we have explored the data table, we generate a clean version of the table.

```{r}
install.packages("lubridate")
library(lubridate)
```

```{r}
cleaned_data<- as.data.frame(matrix(0, ncol = 0, nrow = nrow(data_frame)))

cleaned_data$EntryDateTime=mdy_hm(data_frame$EntryDateTime)
cleaned_data$EntryDay = wday(cleaned_data$EntryDateTime, label=TRUE)
cleaned_data$EntryMonthDay = mday(cleaned_data$EntryDateTime)
cleaned_data$EntryYearDay = yday(cleaned_data$EntryDateTime)

cleaned_data$EntryMonth = month(cleaned_data$EntryDateTime, label=TRUE)
cleaned_data$EntryYear = year(cleaned_data$EntryDateTime)
cleaned_data$EntryHour = hour(cleaned_data$EntryDateTime)

cleaned_data$ExitDateTime=mdy_hm(data_frame$ExitDateTime)
cleaned_data$ExitDay = wday(cleaned_data$ExitDateTime, label=TRUE)
cleaned_data$ExitMonthDay = mday(cleaned_data$ExitDateTime)
cleaned_data$ExitYearDay = yday(cleaned_data$ExitDateTime)

cleaned_data$ExitMonth = month(cleaned_data$ExitDateTime, label =TRUE)
cleaned_data$ExitYear = year(cleaned_data$ExitDateTime)
cleaned_data$ExitHour = hour(cleaned_data$ExitDateTime)

cleaned_data$LastPayDateTime = mdy_hm(data_frame$LastPayDateTime)
cleaned_data$TimeSpent_Hours = round((cleaned_data$ExitDateTime-cleaned_data$EntryDateTime)/3600,digits=2)
cleaned_data$TimeSpentExiting_Hours = round((cleaned_data$ExitDateTime-cleaned_data$LastPayDateTime)/3600,digits=2)
cleaned_data$SiteCarParkNumber =data_frame$SiteCarParkNumber
cleaned_data$EntryDeviceNumber =data_frame$EntryDeviceNumber
cleaned_data$ExitDeviceNumber = data_frame$ExitDeviceNumber
cleaned_data$PaymentNumber =data_frame$PaymentNumber
cleaned_data$CreditCardNumber =data_frame$CreditCardNumber
cleaned_data$TicketType = data_frame$TicketType

distinct_txns = distinct(cleaned_data)
```


####Group by credit card number where a unique visitor is a credit card number.


####Subset by Parking Garage

```{r}
cleaned_data_dlt_rows = filter(cleaned_data,abs(TimeSpent_Hours)<200)
cleaned_data_dlt_rows2 =filter(cleaned_data_dlt_rows, abs(TimeSpentExiting_Hours)<200)
View(cleaned_data_dlt_rows2)
creditcardnumbers = group_by(cleaned_data_dlt_rows, CreditCardNumber)
uniquevisitors = summarise(creditcardnumbers, avg_TimeSpent_Hours = round(mean(TimeSpent_Hours),digits=2),
                           min_TimeSpent_Hours = min(TimeSpent_Hours),
                           max_TimeSpent_Hours = max(TimeSpent_Hours),
                           total_TimeSpent_Hours = sum(TimeSpent_Hours),
                           total_txns = n())
View(uniquevisitors)
```

####hourly/daily exits
```{r}
hourly_exits = group_by(cleaned_data_dlt_rows, ExitDeviceNumber, ExitDay, ExitHour)
total_hourly_exits = summarise(hourly_exits,
                           count_hourly_exits = n())
View(total_hourly_exits)
plot(total_hourly_exits)
```

####hourly/daily entries
```{r}
hourly_entries = group_by(cleaned_data_dlt_rows, EntryDeviceNumber, EntryDay, EntryHour)
total_hourly_entries = summarise(hourly_entries,
                               count_hourly_entries = n())
View(total_hourly_entries)
plot(total_hourly_entries)
```